How are galaxies assigned to halos? \\ Searching for assembly bias in the SDSS galaxy clustering

### 2017-03-06 text/ms.tex
Ref. Report Minor #5 addressed; Minor #7 partly addressed

In the $\mathtt{SMDP}$ simulation, as discussed in \citet{halodemographic}, the Rockstar algorithm can reliably resolve halos with $\geq 100$
%%% CHH: grammar error
{\bf \color{dred} particles, which corresponds to $M_{\rm vir} \geq 9.6 \times 10^{9} \; h^{-1}M_{\odot}$.}
%%%
%The advantage of using the $\mathtt{SMDP}$ simulation is that it satisfies both the size and resolution 
%requirements of studying the galaxy--halo connection in a wide range of luminosity thresholds. 
%%%
%%% Re-worded 
{\bf \color{dred}
    The $\mathtt{SMDP}$ simulation provides a number of advantages by satisfying both the size 
    and resolution requirements of studying the galaxy--halo connection in a wide range of 
    luminosity thresholds. For fainter galaxy samples, the faintest galaxies reside in lower 
    mass halos, which requires high resolution. Meanwhile for luminous galaxy samples, their 
    lower number densities requires a large comoving volume. 
}
%%%
Furthermore, since we are studying the higher order halo occupation statistics, the concentration-dependence in particular, it is important to use a simulation that can resolve the internal structure of halos. 
%%% CHH: see issue in github
{\bf \color{darkgreen}
    In the context of Subhalo-Abundance matching modeling accuracy of which depends on subhalo completeness especially in the low mass limit, the $\mathtt{SMDP}$ simulation has been used to model the faintest galaxy samples in the SDSS data~\citep[see][]{hod_vs_sham}. 
}
%%%
The added advantage of using the $\mathtt{SMDP}$ simulation over some of the other industrial simulation boxes commonly used in the literature such as $\mathtt{Bolshoi}$ (\citealt{Klypin2011,smallmultidark}) simulation is its larger comoving volume. Larger volume makes this simulation more suitable for performing inference with $L_{\star}$ (corresponding to $M_{\rm r} \sim -20.44$, see \citealt{blanton2003}) and more luminous than $L_{\star}$ galaxy samples that occupy larger comoving volumes.

@@

For populating the halos with galaxies, we follow the procedure described in \citet{2016arXiv160701782H}, and \citet{decorated}. The central galaxies are assumed to be at the center of the host dark matter halos. We assume that the central galaxies are at rest with respect to the bulk motion of the halos and their velocities are given by the velocity of the center of mass of their host halo. Note that this assumption is shown to be violated in brighter than $L_{\star}$ galaxy samples (see \citealt{guo2015}). But since we are not considering the redshift space 2PCF multipoles in our study, we do not expect this velocity bias to impact our inference. We place the satellite galaxies within the virial radius of the halo following a Navarro-Frenk-White profile (hereafter NFW; \citealt{nfw}). This approach is different from other simulation-based halo occupation modeling techniques (see \citealt{hod_vs_sham,zheng_guo}) in that the positions of the satellites are not assigned to the dark matter particles in the $N$-body simulation. 
The concentration of the NFW profile is given by the empirical mass-concentration relation provided by \citet{nfw_c(M)}. The velocities of the satellite galaxies are given by two components. The first component is the velocity of the host halo. The second component is the velocity of the satellite galaxy with respect to the host halo which is computed following the solution to the NFW profile Jeans equations (\citealt{more2010}). We refer the readers to \citet{decorated} for a more comprehensive and detailed discussion of the forward modeling of the galaxy mock catalogs.

@@

\subsection{constraints and interpretations}

Our constraints on the assembly bias parameters fall into two main categories. First, the satellite 
assembly bias parameter $\asat$ and the second, the central assembly bias parameter $\acen$. As shown in Figure \ref{fig:bias}, for all the eight luminosity-threshold samples in the SDSS DR7 data, our constraints on the parameter $\asat$ are 
consistent with zero. On the other hand, our constraints on the parameter $\acen$---albeit not tightly constrained---show a trend which can be summarized as the following. In the most luminous galaxy samples, i.e. $M_{\rm r} < -21.5$ and $M_{\rm r} < -21$, $\acen$ is poorly constrained and the constraints are equivalent to zero. As we investigate less luminous samples, $M_{\rm r} < -20.5 , -20, -19.5$, our constraints on $\acen$ shift toward positive values, with the $M_{\rm r}<-20$ sample favoring the highest values for $\acen$. Furthermore, the posterior constraints on the assembly bias parameters in the slightly fainter samples, i.e. $M_{\rm r} < -19$ and i.e. $M_{\rm r} < -18.5$, are consistent with zero. In the faintest galaxy sample, i.e. $M_{\rm r} < -18$, our constraints favor negative values of $\acen$.

@@

A few galaxy--halo connection methods have been proposed in the literature that give rise to assembly bias in the galaxy population. \citet{arz2014} demonstrates that the abundance matching techniques based on $V_{\rm max }$ (as introduced in \citealt{hw2013,reddick2013}) exhibit some levels of assembly bias. 
We aim to provide a comparison between the impact of assembly bias on clustering in these mock catalogs and the mock catalogs predicted from our constraints on the decorated HOD model for $L_{\star}$-type galaxies. In particular, we consider the abundance matching catalogs produced by \citet{hw2013}. These catalogs have been extensively studied for examining potential systematic effects of galaxy assembly bias on cosmological (\citealt{edHOD-weinberg}) and halo occupation (\citealt{arz2014}) parameter inferences.

>>>>

In the $\mathtt{SMDP}$ simulation, as discussed in \citet{halodemographic}, the Rockstar algorithm can reliably resolve halos with $\geq 100$
%%% CHH: grammar error
{\bf \color{dred} particles, which corresponds to $M_{\rm vir} \geq 9.6 \times 10^{9} \; h^{-1}M_{\odot}$.}
%%%
%The advantage of using the $\mathtt{SMDP}$ simulation is that it satisfies both the size and resolution 
%requirements of studying the galaxy--halo connection in a wide range of luminosity thresholds. 
%%%
%%% Re-worded 
{\bf \color{dred}
    The $\mathtt{SMDP}$ simulation provides a number of advantages by satisfying both the size 
    and resolution requirements of studying the galaxy--halo connection in a wide range of 
    luminosity thresholds. For fainter galaxy samples, the faintest galaxies reside in lower 
    mass halos, which requires high resolution. Meanwhile for luminous galaxy samples, their 
    lower number densities requires a large comoving volume. 
}
%%%
Furthermore, since we are studying the higher order halo occupation statistics, the concentration-dependence in particular, it is important to use a simulation that can resolve the internal structure of halos. 
%%% CHH: Re-worded based on (github issue #1)
{\bf \color{dred}
    In the context of Subhalo-Abundance Matching models, which requires subhalo completeness 
    in the low mass limit, the $\mathtt{SMDP}$ simulation has been used to model the faintest 
    galaxy samples in the SDSS data~\citep[see][]{hod_vs_sham}. 
}
%%%
The added advantage of using the $\mathtt{SMDP}$ simulation over some of the other industrial simulation boxes commonly used in the literature such as $\mathtt{Bolshoi}$ (\citealt{Klypin2011,smallmultidark}) simulation is its larger comoving volume. Larger volume makes this simulation more suitable for performing inference with $L_{\star}$ (corresponding to $M_{\rm r} \sim -20.44$, see \citealt{blanton2003}) and more luminous than $L_{\star}$ galaxy samples that occupy larger comoving volumes.

@@

For populating the halos with galaxies, we follow the procedure described in \citet{2016arXiv160701782H}, and \citet{decorated}. The central galaxies are assumed to be at the center of the host dark matter halos. We assume that the central galaxies are at rest with respect to the bulk motion of the halos and their velocities are given by the velocity of the center of mass of their host halo. Note that this assumption is shown to be violated in brighter than $L_{\star}$ galaxy samples (see \citealt{guo2015}). But since we are not considering the redshift space 2PCF multipoles in our study, we do not expect this velocity bias to impact our inference. We place the satellite galaxies within the virial radius of the halo following a Navarro-Frenk-White profile (hereafter NFW; \citealt{nfw}). This approach is different from other simulation-based halo occupation modeling techniques (see \citealt{hod_vs_sham,zheng_guo}) in that the positions of the satellites are not assigned to the dark matter particles in the $N$-body simulation. 
%%% Missing citation
{\bf \color{darkgreen} 
    The concentration of the NFW profile is given by the empirical 
    mass-concentration relation provided by \citet{nfw_c(M)}. 
}
%%%
The velocities of the satellite galaxies are given by two components. The first component is the velocity of the host halo. The second component is the velocity of the satellite galaxy with respect to the host halo which is computed following the solution to the NFW profile Jeans equations (\citealt{more2010}). We refer the readers to \citet{decorated} for a more comprehensive and detailed discussion of the forward modeling of the galaxy mock catalogs.

@@

\subsection{{\color{dred} Constraints and Interpretations}}
%Our constraints on the assembly bias parameters fall into two main categories. First, the satellite 
%assembly bias parameter $\asat$ and the second, the central assembly bias parameter $\acen$. 
%%% Re-worded
{\bf \color{dred}  
    In this section, we present the constraints derived for the two assembly 
    bias parameters: the satellite assembly bias parameter ($\asat$) and the 
    central assembly bias parameter $\acen$. 
} 
%%%
As shown in Figure \ref{fig:bias}, for all the eight luminosity-threshold samples in the SDSS DR7 data, our constraints on the parameter $\asat$ are 
consistent with zero. On the other hand, our constraints on the parameter $\acen$--- albeit not tightly constrained--- show a trend which can be summarized as the following. In the most luminous galaxy samples, i.e. $M_{\rm r} < -21.5$ and $M_{\rm r} < -21$, $\acen$ is poorly constrained and the constraints are equivalent to zero. As we investigate less luminous samples, $M_{\rm r} < -20.5 , -20, -19.5$, our constraints on $\acen$ shift toward positive values, with the $M_{\rm r}<-20$ sample favoring the highest values for $\acen$. Furthermore, the posterior constraints on the assembly bias parameters in the slightly fainter samples, i.e. $M_{\rm r} < -19$ and i.e. $M_{\rm r} < -18.5$, are consistent with zero. In the faintest galaxy sample, i.e. $M_{\rm r} < -18$, our constraints favor negative values of $\acen$.

@@

A few galaxy--halo connection methods have been proposed in the literature that give rise to assembly bias in the galaxy population. 
%%%% CHH: (Ref. Report Minor #5) re-worded and Conroy et al. (2006) included in citations. 
{\bf \color{dred} 
    \citet{arz2014} demonstrates that the abundance matching techniques based on $V_{\rm max }$ 
    \citep{conroy:2006aa,hw2013,reddick2013} exhibit some levels of assembly bias. 
}
We aim to provide a comparison between the impact of assembly bias on clustering in these mock catalogs and the mock catalogs predicted from our constraints on the decorated HOD model for $L_{\star}$-type galaxies. In particular, we consider the abundance matching catalogs produced by \citet{hw2013}. These catalogs have been extensively studied for examining potential systematic effects of galaxy assembly bias on cosmological (\citealt{edHOD-weinberg}) and halo occupation (\citealt{arz2014}) parameter inferences.
